### 2022-11-28 Secondo Anno/Analisi Matematica 2/mat.tex
introduzione alle serie

\documentclass{book}
\usepackage[a4paper,top=2.0cm,bottom=2.0cm,left=2.0cm,right=3.0cm]{geometry}

%\documentclass[pdftex,10pt,a4paper]{book}
%\usepackage[paperwidth=19cm,
%paperheight=26cm, outer=2cm, 
%top=1.5cm, bottom=1.5cm]{ geometry}

\usepackage[english,italian]{babel} %l'ultima lingua è quella che legge per i titoli
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc,url}
\usepackage{titlesec}
\usepackage{easylist}
\usepackage{hanging}

\usepackage[pdftex,colorlinks]{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=black,
	filecolor=magenta,
	urlcolor=cyan,
}
\usepackage{hypcap}
\usepackage{blindtext}
\usepackage{tipa}
\usepackage{epigraph}
\usepackage{enumerate}
\usepackage{longtable}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{pbox}
\usepackage{fancyhdr}
\usepackage{cancel}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{subfig}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{textcomp}
\usepackage{circuitikz}
\usepackage{pifont}
\usepackage{imakeidx}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{listings}
\usepackage{color}
\usepackage{upgreek}
\usepackage{tasks}
\usepackage{exsheets}
\usepackage{pgfplots}
\usepackage{amsthm}
\usepackage{wasysym}

\usepackage{showframe}
\renewcommand\ShowFrameLinethickness{0.15pt}
%\renewcommand*\ShowFrameColor{\color{red}}

%\usepackage{showkeys} %serve per mostrare le etichette "tag" o target, va tolta per la versione definitiva;

\SetupExSheets[question]{type=exam}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ 
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  firstnumber=1000,                % start line enumeration with line 1000
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Octave,                 % the language of the code
  morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

% definizioni
\newtheorem{esercizio}{Esercizio}
\newtheorem{teorema}{Teorema}
\newtheorem{nota}{Nota}
\newtheorem{defi}{Definizione}
\newtheorem{esempio}{Esempio}
\newtheorem{svol}{Svolgimento}
\newtheorem{pro}{Proprietà}
\linespread{1.2} % l'interlinea

\frenchspacing

\newcommand{\abs}[1]{\lvert#1\rvert}

\usepackage{floatflt,epsfig}

\usepackage{multicol}
\newcommand\yellowbigsqcup[1][\displaystyle]{%
  \fboxrule0pt
  \ifx#1\textstyle\fboxsep-0.6pt\else\fboxsep-1.25pt\fi
  \mathrel{\fcolorbox{white}{yellow}{$#1\bigsqcup$}}}

\title{Appunti di Matematica}
\author{Nicola Ferru}
\date{}
\makeindex[columns=3, title=Alphabetical Index, intoc]

\begin{document}
%\maketitle
\include{./title/titlepage.tex}

\tableofcontents
%\listoftables
\listoffigures

\input{pag/premesse.tex}
\input{pag/simboli.tex}
\input{pag/intro.tex}
\input{pag/derivate_parziali.tex}
\input{pag/differenzialita.tex}
\input{pag/integrali.tex}
\chapter{Successioni e serie}
\section{Successioni di costanti}
\begin{defi}
  Si definisce successione $a_n$ una funzione in N (numeri naturali) e codominio in R.\\
  È possibile rappresentare una successione sul pieno cartesiano, con l'asse delle
  ascisse N e quello delle ordinate R.
\end{defi}

\section{Limite di successioni}
\subsection{Limite finito di una successione}
\begin{defi}
  Si definisce limite finito della succesione $a_n$ ($a_n$ converge) il numero reale
  $\gamma$ tale che:
  \begin{eqnarray*}
    \lim_{n\to \infty} a_n=\gamma & \forall \xi >0, \exists P_\xi \in N:\forall n > P_\xi\\
                                  & \abs{a_n-\gamma} < \xi
  \end{eqnarray*}
  Una successione che per $n\to \infty$ ha limite finito si dice {\color{red}convergente}
\end{defi}
\subsubsection{Limite infinito di una successione}
$bn\to +\infty$ se
\begin{eqnarray*}
  \lim_{n\to\infty}bn=+\infty & \forall M>0, \exists P_M:\forall n > P_m & b_R>M
\end{eqnarray*}
Una successione che per $n\to \infty$ ha infinito si dice {\color{red}divergente} --
Se una successione nè converge, nè diverge, diciamo che è {\color{red}irregolare} o
{\color{red}indeterminata}, cioè non esiste $\lim\limits_{n\to\infty}$.
\section{Successioni Limitatre, illimitate, crescenti e decrescenti}
Limitata superiormente
\begin{eqnarray*}
  a_n \text{ è {\color{red} limitata superiormente} se } & \exists k:\forall n \in N
  & a_n\leq k\\
  a_n \text{ è {\color{red} limitata inferiormente} se} & \exists h: \forall n\in N
  & a_n \geq h\\
  a_n \text{ è {\color{red} limitata} se} & \text{è limita inferiormente e superiormente}
  & h\leq a_n\leq k\\
  a_n \text{ è {\color{red}monotona crescente} se} & \forall n: & a_n < a_{n+1}\\
  a_n \text{ è {\color{red}monotona decrescente} se} & \forall n: & a_n > a_{n+1}
\end{eqnarray*}

\printindex
\end{document}