### 2022-12-15 Primo anno/Fisica/fisica.tex
aggiunte fisica 1

\documentclass{book}
\usepackage[a4paper,top=2.0cm,bottom=2.0cm,left=2.0cm,right=3.0cm]{geometry}

%\documentclass[pdftex,10pt,a4paper]{book}
%\usepackage[paperwidth=19cm,
%paperheight=26cm, outer=2cm, 
%top=1.5cm, bottom=1.5cm]{ geometry}

\usepackage[english,italian]{babel} %l'ultima lingua è quella che legge per i titoli
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc,url}
\usepackage{titlesec}
\usepackage{easylist}
\usepackage{hanging}

\usepackage[pdftex,colorlinks]{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=black,
	filecolor=magenta,
	urlcolor=cyan,
}
\usepackage{hypcap}
\usepackage{blindtext}
\usepackage{tipa}
\usepackage{epigraph}
\usepackage{enumerate}
\usepackage{longtable}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{pbox}
\usepackage{fancyhdr}
\usepackage{cancel}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{subfig}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{textcomp}
\usepackage{circuitikz}
\usepackage{pifont}
\usepackage{imakeidx}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{listings}
\usepackage{color}
\usepackage{upgreek}
\usepackage{tasks}
\usepackage{exsheets}
\usepackage{pgfplots}
\usepackage{amsthm}
\usepackage{wasysym}

\usepackage{showframe}
\renewcommand\ShowFrameLinethickness{0.15pt}
%\renewcommand*\ShowFrameColor{\color{red}}

%\usepackage{showkeys} %serve per mostrare le etichette "tag" o target, va tolta per la versione definitiva;

\SetupExSheets[question]{type=exam}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ 
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  firstnumber=1000,                % start line enumeration with line 1000
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Octave,                 % the language of the code
  morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

% definizioni
\newtheorem{esercizio}{Esercizio}
\newtheorem{teo}{Teorema}
\newtheorem{nota}{Nota}
\newtheorem{defi}{Definizione}
\newtheorem{esempio}{Esempio}
\newtheorem{svol}{Svolgimento}
\newtheorem{pro}{Proprietà}
\newtheorem{sol}{soluzione}
\linespread{1.2} % l'interlinea

\frenchspacing

\newcommand{\abs}[1]{\lvert#1\rvert}

\usepackage{floatflt,epsfig}

\usepackage{multicol}
\newcommand\yellowbigsqcup[1][\displaystyle]{%
  \fboxrule0pt
  \ifx#1\textstyle\fboxsep-0.6pt\else\fboxsep-1.25pt\fi
  \mathrel{\fcolorbox{white}{yellow}{$#1\bigsqcup$}}}

\title{Appunti Fisica 1 \& 2}
\author{Nicola Ferru}
\date{}

\begin{document}
%\maketitle
\include{./title/titlepage.tex}
\tableofcontents
\listoftables
\listoffigures
% pagine dedicate
\input{pag/premesse.tex}
\input{pag/simboli.tex}
\part{fisica 1}
\input{pag/unita di misura.tex}
\input{pag/energia.tex}
\chapter{I moti}
\section{Moto di Roto-Traslario}
Analizziamo il moto più generale di un sistema meccanico isolato, cioè privo di forze
esterne.
\begin{eqnarray*}
  M\vec{a}_{CM}=\displaystyle\sum \vec{F}_{ext}=0 & \begin{matrix}
                                                      \text{In assenza di forze esterne
                                                      il centro di massa si muove di
                                                      moto retilineo}\\
                                                      \text{univorme oppure sta
                                                      fermo (in un sistema di
                                                      riferimento inerziale)}
                                                    \end{matrix}
\end{eqnarray*}
Posiamo studiare quuesto moto rispetto ad un sistema inerziale fermo dal quale
osservare il moto del corpo rigido:
\begin{figure}[ht]
  \centering
  \includegraphics[width=8cm]{img/finiti/roto-trasl.eps}
  \caption{Sistema di riferimento per il moto Roto-Traslario}
  \label{fig:roto-traslario}
\end{figure}
Nel sistema $o^\prime x^\prime y^\prime$ nell'energia cinetica sarà completamente
rotazionale e quindi data da:
\begin{eqnarray*}
  I_{cm}=\int r^{\prime2} dm & \begin{matrix}
                                 \text{dove $I_{cm}$ il momento d'inezia del corpo}\\
                                 \text{per la rotazione attorno al \it cm.}
                               \end{matrix}
\end{eqnarray*}
Nel sistema $oxy$, $\vec{r}^\prime$ può essere espresso come:
\begin{eqnarray*}
  \vec{r}=\vec{r}^\prime+\vec{r}_{cm} & \to & \vec{v}=\vec{v}^\prime+\vec{v}_{cm}
\end{eqnarray*}
Scriviamo quindi l'energia cinetica del corpo rigido nel sistema di riferimento $oxy$:
\begin{equation}
  K=\frac{1}{2}\displaystyle\sum \underbrace{m_iv^2\to K}_{\text{corpo continuo}}=
  \frac{1}{2}\int v^2 dm
\end{equation}
\input{pag/modelli atomici.tex}
\input{pag/fisica2.tex}
\end{document}